---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Birdr

<!-- badges: start -->
<!-- badges: end -->

The goal of Birdr is to retrieve select data from the eBird api and create maps/graphics that could be useful to birders in the United States

## Installation

You can install the the development version from [GitHub](https://github.com/) with:

```{r eval = FALSE}
# install.packages("devtools")
devtools::install_github("sjmarks/Birdr")

```

```{r}
library(Birdr)
```

## eBird keys

Most functions in Birdr require an API key

Please obtain a key here before attempting to use the package - (https://ebird.org/api/keygen)

## Obtain United States region codes

Certain functions in Birdr require you to specify a region in the US to look at/visualize data from

You can use the function `pick_UScode()` to determine what you need to input for region specifying arguments.

The function returns a subnational1 regional code for the US by sepcifying the state of interest:

```{r}
pick_UScode(state = "California", ebirdkey = '1ls8sv5baidn')
```

OR a subnational2 regional code for the US if a county name is supplied:

```{r}
pick_UScode(state = "California", county = "San Luis Obispo", ebirdkey = '1ls8sv5baidn')
```


## GET recent notable obs for a region

You can use the function `get_notable_obs()` to retrieve recent notable bird observations from eBird AI for a specified US region:

```{r}
get_notable_obs(region_code = "US-CA-079", ebirdkey = '1ls8sv5baidn', back = 2)
```

## Create a chart of recent obs counts for a region

You can create a dot plot showing counts of recent notable birds for a specified region using `get_notable_chart()`:

```{r}
get_notable_chart(region_code = "US-CA-079", ebirdkey = '1ls8sv5baidn', obs_to_plot = 25, back = 15)
```

## Create a map of local hotspots

You can create a map of local birding hotspots relevent to your location with `map_local_hotspots()`: 

```{r}
map_local_hotspots(latitude = 35.28, longitude = -120.65, key = '1ls8sv5baidn')
```

## GET a dataframe of all the bird species on eBird

You can get a dataframe of all birds on eBird with `get_all_birds()`:

```{r}
get_all_birds(key = '1ls8sv5baidn') %>% 
  head(5)
```

## GET a dataframe of all local hotspots or hotspots for a region

You can get both local and regional hotspots with these two functions `get_local_hotspots()` and `get_hotspots()`:

```{r}
get_local_hotspots(35.28, -120.65, key = '1ls8sv5baidn')
```
```{r}
get_hotspots("US-CA-079", key = '1ls8sv5baidn')
```

## GET a dataframe of regions 

You can get a dataframe of regions (countries, state level, or county level) with `get_regions()`: 

```{r}
get_regions(region_type = "subnational2", parent_region = "US-CA", key = '1ls8sv5baidn')
```

